BioTime

### 2019-04-25 biotime.Rmd
Tried to use synonyms() to check for id's with other providers, but it doesn't return the input column so you can't join back to original data.

#Getting matches by sci name and common name 
biotime_ids <- by_name(unique(data$genus_species), "itis") %>%
  bind_rows(by_common(unique(data$genus_species), "itis")) %>%
  filter(taxonRank == "species") %>% #only want ID's to species, since that's the level of the trait data
  drop_na(acceptedNameUsageID)
```

@@

@@

Join bird ids back to original trait data (to get data that didn't have an id match)
```{r}
#only want one column for old name and one column for new name, these are the old ID columns to remove
col_ex <- c("passnonpass", "iocorder", "blfamilylatin", "blfamilyenglish", "blfamsequid", "taxo")

@@

Do the same for mammal trait data
```{r}
mamm_trait_data <- mammal_ids %>%
  select(id = acceptedNameUsageID, input, scientificName) %>%
  distinct() %>%
  right_join(mamm_trait %>% select(-mswfamilylatin), by = c("input" = "scientific"), na_matches = "never") %>%
  drop_na(id) %>%
  distinct() 
```

check if all bird biotime data w/id's has an elton match. There are 38 species without a match
```{r}
biotime_data %>%
  filter(taxa == "Birds") %>%
  select(id) %>%
  distinct() %>%
  filter(!id %in% bird_trait_data$id) %>%
  dim()
```

>>>>

#Getting matches by sci name and common name 
biotime_ids <- by_name(unique(data$genus_species), "itis") %>%
  bind_rows(by_common(unique(data$genus_species), "itis")) %>%
  #filter(taxonRank == "species") %>% #only want ID's to species, since that's the level of the trait data
  drop_na(acceptedNameUsageID)
```

@@

Do any unmatched species have matches with other providers?
```{r}
get_match_counts(biotime_data %>% filter(is.na(id)), "matchingName")

#match all the un-ID'd names to other providers, check if synonyms given by those providers match known ITIS species 
syn_res <- bind_rows(synonyms(biotime_data %>% filter(is.na(id)) %>% pull(matchingName) %>% unique(), "ott"),
                  synonyms(biotime_data %>% filter(is.na(id)) %>% pull(matchingName) %>% unique(), "gbif"),
                  synonyms(biotime_data %>% filter(is.na(id)) %>% pull(matchingName) %>% unique(), "col"),
                  synonyms(biotime_data %>% filter(is.na(id)) %>% pull(matchingName) %>% unique(), "ncbi"),
                  synonyms(biotime_data %>% filter(is.na(id)) %>% pull(matchingName) %>% unique(), "wd"),
                  synonyms(biotime_data %>% filter(is.na(id)) %>% pull(matchingName) %>% unique(), "iucn")
                  ) %>%
  pull(synonym) %>% 
  na.omit() %>% 
  unique() %>%
  by_name("itis") %>%
  select(id = acceptedNameUsageID, input, scientificName) %>% 
  drop_na(id) %>% 
  left_join(data %>% select(-species, -genus), by = c("input" = "genus_species")) %>%
  rename(matchingName = input)

#Below add's more rows, should just replace them 
biotime_data <- biotime_data %>% filter(!matchingName %in% syn_res$matchingName) %>%
  bind_rows(syn_res)

#These two should be equal, why aren't they??
dim(syn_res)  
biotime_data %>% filter(matchingName %in% syn_res$matchingName) %>% dim()
biotime_data %>% filter(is.na(id)) %>% dim()
```

@@

Join bird ids back to original trait data
```{r}
#only want one column for old name and one column for new name, these are the old ID columns to remove
col_ex <- c("passnonpass", "iocorder", "blfamilylatin", "blfamilyenglish", "blfamsequid", "taxo")

@@

Do the same for mammal trait data
```{r}
mamm_trait_data <- mammal_ids %>%
  select(id = acceptedNameUsageID, input, scientificName) %>%
  right_join(mamm_trait %>% select(-mswfamilylatin), by = c("input" = "scientific"), na_matches = "never") %>%
  drop_na(id) %>%
  distinct() 
```

#Checking Coverage of ID's and trait data -> biotime

check if all bird & mammal biotime data w/id's has an elton match. There are 258 species without a match (38 are birds).
```{r}
biotime_data %>%
  #filter(taxa == "Birds") %>%
  select(id) %>%
  distinct() %>%
  filter(!id %in% bird_trait_data$id) %>%
  dim()
```
